---
title: "BSED scenarios"
author: "Renata Diaz"
date: "5/28/2019"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(replicatebecs)
```

This is what BSEDs would look like, and what the results of a bootstrapping analysis would be, under the assumptions of a) a uniform size-abundance distribution ([Uniform size-abundance]) and b) a uniform size-energy ([Uniform size-energy])distribution. 

Using the Andrews LTER small mammal data for demonstration.

Load the Andrews data, generate BSED, and get summary variables:
```{r load andrews}
andrews <- load_paper_data()[[1]]
head(andrews)

andrews_bsed <- andrews %>%
  add_energy_sizeclass() %>%
  make_bsed()
```

```{r andrews descriptor variables}
andrews_vars <- list(nind = nrow(andrews),
                    nsp = length(unique(andrews$individual_species_ids)),
                    min_size = min(andrews$individual_sizes),
                    max_size = max(andrews$individual_sizes),
                    nsize_classes = nrow(andrews_bsed),
                    size_classes = andrews_bsed$size_class,
                    total_energy = sum((andrews$individual_sizes)^(3/4)))
```

## Uniform size-abundance

For a uniform size-abundance distribution, individuals' sizes are uniformly distributed from the minimum to the maximum body size in the community.

For simulation, this translates as:

```{r uniform size abund sampling}
unif_sa_sample <- runif(n = andrews_vars$nind, min = andrews_vars$min_size, max = andrews_vars$max_size)
```

We use this vector of masses to generate a BSED for the simulated community:

```{r unif size abund bsed}
unif_sa_bsed <- andrews %>%
  dplyr::mutate(individual_sizes = unif_sa_sample) %>% # Change measured masses to simulated masses
  add_energy_sizeclass() %>%
  make_bsed()

head(unif_sa_bsed)
```

To define the "true" form of the distribution, we can define a "community" with exactly one individual of every size from the minimum to the maximum in the community and generate the corresponding BSED. `replicatebecs::calculate_uniform_size_abund_bsed` does this. 

```{r calculate unif size abund bsed}
unif_sa_calculated <- calculate_uniform_size_abund_bsed(andrews)
```


```{r plot unif size abund bseds, fig.width = 15, fig.height = 5, echo = F}
andrews_plot = plot_bsed(andrews_bsed, bsed_name = "Real Andrews")

unif_sa_sampled_plot = plot_bsed(unif_sa_bsed, bsed_name = "Sampled uniform size-abundance")

unif_sa_calc_plot = plot_bsed(unif_sa_calculated, bsed_name = "'True' uniform size-abundance")

unif_sa_plots = gridExtra::grid.arrange(andrews_plot, unif_sa_sampled_plot, unif_sa_calc_plot, nrow = 1)

invisible(unif_sa_plots)

```


## Uniform size-energy

For a uniform size-energy distribution, all body size classes have an equal share of the energy use in the community.

This corresponds to a steeply negative relationship between size and abundance because a) larger individuals use more energy per capita and b) the body size classes are binned in log units, so the larger body size classes have the same amount of energy divided across a broader portion of the size spectrum. 

To sample:

```{r uniform size energy sample}

unif_se_bsed <- andrews_bsed %>%
  dplyr::mutate(total_energy = runif(n = andrews_vars$nsize_classes, min = 0, max = andrews_vars$total_energy)) %>% # the maximum doesn't matter, because we will convert it to proportional
  dplyr::mutate(total_energy_proportional = total_energy / sum(total_energy))

```


To calculate:

```{r uniform size energy calculation}
unif_se_calculated = andrews_bsed %>%
  dplyr::mutate(total_energy_proportional = 1/andrews_vars$nsize_classes)
```


```{r plot unif size energy bseds, fig.width = 15, fig.height = 5, echo = F}
andrews_plot = plot_bsed(andrews_bsed, bsed_name = "Real Andrews")

unif_se_sampled_plot = plot_bsed(unif_se_bsed, bsed_name = "Sampled uniform size-energy")

unif_se_calc_plot = plot_bsed(unif_se_calculated, bsed_name = "'True' uniform size-energy")

unif_se_plots = gridExtra::grid.arrange(andrews_plot, unif_se_sampled_plot, unif_se_calc_plot, nrow = 1)

invisible(unif_se_plots)

```


