---
title: "Replication of Ernest (2005)"
author: "Renata Diaz"
date: "5/4/2019"
output: github_document
---

```{r setup}
library(revisitbecs)
library(ggplot2)
```

# Import data

```{r import data}

process_raw_data()

data_files <- list.files(path = paste0(here::here(), '/data/paper/processed'), full.names = T)

communities <- list()

for(i in 1:length(data_files)) {
communities[[i]] <- read.csv(data_files[[i]], stringsAsFactors = F)
}

rm(data_files)
rm(i)

ncommunities = length(communities)

```


We need to 1) estimate each individual's energy use (as $m^{3/4}$ where $m$ is mass in g) and 2) assign each individual to a size class. We will divide the communities into size classes of .2 log units. 

```{r make community tables}
community_tables <- list()

for(i in 1:length(communities)){
  community_tables[[i]] <- make_community_table(community = communities[[i]])
}

```


# Construct BSEDs for those communities

Now we can construct a body size-energy distribution for each community. This is a summary, for each size class, of how much energy all the individuals (regardless of species ID) in that size class use. We will work with these distributions standardized according to the total energy used by the whole community. 

```{r make real bseds}

real_bseds <- list()

for(i in 1:ncommunities){
  real_bseds[[i]] <- make_bsed(community_tables[[i]], decimals = 1)
}

```

# BSED plots 

```{r plot real BSEDS, fig.width = 6, fig.height = 4, echo = F}


for(i in 1:ncommunities) {
  bsed_plot = ggplot(data = real_bseds[[i]], aes(x = size_class, y = total_energy_proportional)) +
    geom_bar(stat = 'identity', aes(x = as.factor(real_bseds[[i]]$size_class_g), y = real_bseds[[i]]$total_energy_proportional)) +
    ggtitle(paste0("BSED for community ", i)) +
    theme_bw()
  
  print(bsed_plot)
  
}

```


# Calculate and plot energetic dominance

```{r energetic dominance for real communities}

dominance_values <- vector(mode = "numeric")

for(i in 1:ncommunities) {
  these_modes <- energetic_dominance(community_tables[[i]])
  
  these_modes <- these_modes %>%
    dplyr::select(mode_id, e_dominance) %>%
    dplyr::distinct()
  
  dominance_values <- c(dominance_values, these_modes$e_dominance)
}

anyNA(dominance_values)

dominance_values <- as.data.frame(dominance_values)

e_dominance_plot <- ggplot(data = dominance_values) + 
  geom_histogram(binwidth = 0.1, aes(x = dominance_values)) +
  xlim(-0.1, 1.1) + 
  theme_bw()

e_dominance_plot

```

# Compare each real BSED to 10000 bootstraps (DOI 95% interval)

```{r bootstrap BSEDs compared to uniform}
nsamples = 10000
for(i in 1:ncommunities){ 
  sampled_communities_doi <- replicate(nsamples, boostrap_unif_bsed_doi(communities[[i]]))
  
  real_doi <- doi(real_bseds[[i]]$total_energy_proportional)
  
  p_greater_doi <- length(which(sampled_communities_doi > real_doi)) / nsamples
  
  print(p_greater_doi)
  
}


```

# Compare all pairwise communities BSEDs

```{r bootstrap BSEDs compared to each other}

nsamples = 10000

all_pairs_matrix <- combn(1:ncommunities, m = 2)

p_comparison <- vector(length = ncol(all_pairs_matrix), mode = 'numeric')

for(i in 1:ncol(all_pairs_matrix)) {
  
  first = all_pairs_matrix[1, i]
    second = all_pairs_matrix[2, i]
sampled_pair_doi <- replicate(nsamples, boostrap_crosscomm_bseds(communities[[first]], communities[[second]]))


both_bseds <- real_bseds[[first]] %>%
  dplyr::full_join(real_bseds[[second]], by = c("size_class", "size_class_g")) %>%
  dplyr::mutate(total_energy_proportional.x = replace(total_energy_proportional.x, is.na(total_energy_proportional.x), 0),
                total_energy_proportional.y = replace(total_energy_proportional.y, is.na(total_energy_proportional.y), 0))

real_doi <- doi(both_bseds$total_energy_proportional.x,
                both_bseds$total_energy_proportional.y)

p_greater_doi <- length(which(sampled_pair_doi > real_doi)) / nsamples

p_comparison[i] <- p_greater_doi
}

p_comparison

length(which(p_comparison > 0.05)) / length(p_comparison)

```

# Construct BSDs for real communities

```{r make real BSDs}
real_bsds <- list()
for(i in 1:ncommunities) {
  real_bsds[[i]] <- make_bsd(community_tables[[i]], decimals = 2)
}
```

# Plot them

```{r plot real BSDs} 

for(i in 1:ncommunities){ 
  this_bsd <- real_bsds[[i]]
  bsd_plot <- ggplot(data = this_bsd, aes(x = size_class, y =  n_species_proportional)) +
    geom_point(data = this_bsd, aes(x = as.factor(size_class_g), y= n_species_proportional)) + 
    theme_bw()
  print(bsd_plot)
  }

```

# Compare each real BSD to uniform (d-corrected KS)

```{r compare bsds to uniform}
for (i in 1:ncommunities) {
  this_bsd <- real_bsds[[i]]
  this_ks <- ks.test(this_bsd$n_species_proportional, punif)
  print(this_ks$p.value)
}

```


# Compare all pairwise communities BSDs (KS)

```{r pairwise BSD comparisons}

all_pairs_matrix <- combn(1:ncommunities, m = 2)

ks_p_comparison <- vector(length = ncol(all_pairs_matrix), mode = 'numeric')

for(i in 1:ncol(all_pairs_matrix)) {
  
  first = all_pairs_matrix[1, i]
    second = all_pairs_matrix[2, i]

both_bsds <- real_bsds[[first]] %>%
  dplyr::full_join(real_bsds[[second]], by = c("size_class", "size_class_g")) %>%
  dplyr::mutate(n_species_proportional.x = replace(n_species_proportional.x, is.na(n_species_proportional.x), 0),
                n_species_proportional.y = replace(n_species_proportional.y, is.na(n_species_proportional.y), 0))

ks_comparison <- ks.test(both_bsds$n_species_proportional.x,
                both_bsds$n_species_proportional.y)
ks_p_comparison[i] <- ks_comparison$p.value
}

ks_p_comparison

length(which(ks_p_comparison < 0.05))
```